Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year

### 2021-11-04 cover/cover.Rmd
Add Astley paragraph

library(covidcast)
library(epitools)
library(ggplot2)
library(dplyr)
library(purrr)
```

```{r get-data, eval = FALSE}
sources <- c("jhu-csse",
             "doctor-visits",
             "fb-survey",
             "google-symptoms")

@@

start <- as.Date("2020-04-15")
end <- as.Date("2021-10-22")

Select <- function(x, ...) {
  if (nrow(x) == 0) x
  else select(x, ...)
}

x <- covidcast_signals(
  sources, signals,
  start, end,
  geo_type = "state"
) %>% map(function(df) df %>%
            Select(data_source, signal, geo_value, time_value, value))
```

@@

@@

dates <- seq.Date(start, end, by = "1 day")
n_days <- length(dates)

g <- vector(mode = "list", length = length(x))
delta <- 1/500
alpha <- 0.35
xlim <- c(start - 10, end)
ylim <- c(-0.05, 1.05 + 50 * delta)
xp <- xlim[1] - 8
yp <- mean(ylim)

for (i in 1:length(x)) {
  y <- tibble(geo_value = rep(unique(x[[i]]$geo_value), each = n_days),
              time_value = rep(seq.Date(start, end, by = "day"),
                               length(unique(x[[i]]$geo_value))))
  x[[i]] <- full_join(x[[i]], y, by = c("geo_value", "time_value"))

  g[[i]] <- x[[i]] %>%
    filter(geo_value %in% tolower(state.abb)) %>%
    mutate(geo_num = as.integer(as.factor(geo_value))) %>%
    group_by(geo_value) %>%
    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) +
             delta * (geo_num - 1)) %>%
    ungroup() %>%
    ggplot(aes(x = time_value, y = value, group = geo_value)) +
    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
    geom_line(alpha = 0.5, color = colors[i], size = 0.3) +
    theme_void() +
    annotate("text", x = xp, y = yp, label = names[i],
             color = colors[i], angle = 90, size = 3) +
    theme(panel.background = element_rect(fill = "black"))
}

gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
# PNAS wants 21 cm wide by 22.5 cm high
ggsave(file = "cover-graphic.pdf", gg, height = 22.5 / 2.54, width = 21 / 2.54)
```

>>>>

library(covidcast)
library(ggplot2)
library(dplyr)
library(purrr)
```

```{r get-data, eval = FALSE}
sources <- c("usa-facts",
             "doctor-visits",
             "fb-survey",
             "google-symptoms")

@@

start <- as.Date("2020-04-15")
end <- as.Date("2021-10-22")
dates <- seq.Date(start, end, by = "1 day")
geos <- tolower(state.abb)

x <- covidcast_signals(
  sources, signals,
  start, end,
  geo_type = "state", 
  geo_values = geos) %>% 
  map(function(df) df %>%
        select(data_source, signal, geo_value, time_value, value))
```

@@

```{r load-all}
# Inspect missingness: google has lots (privacy threshold)
do.call(rbind, x) %>%
  filter(geo_value %in% tolower(state.abb)) %>%
  group_by(signal, geo_value) %>%
  summarize(nas = length(dates) - n()) %>%
  ungroup() %>% group_by(signal) %>%
  filter(nas > 0)

y <- tibble(geo_value = rep(geos, each = length(dates)),
            time_value = rep(dates, length(geos)))

x[[4]] %>%
  full_join(y, by = c("geo_value", "time_value")) %>%
  group_by(time_value) %>%
  summarize(nas = sum(is.na(value))) %>% 
  filter(nas > 40)
```

@@

g <- vector(mode = "list", length = length(x))
delta <- 1 / length(geos) * 0.1
alpha <- 0.35
line_width <- 0.3
label_size <- 3
xlim <- c(start - 10, end)
ylim <- c(-0.05, 1.05 + length(geos) * delta)
xp <- xlim[1] - 8
yp <- mean(ylim)

for (i in 1:length(x)) {
  g[[i]] <- x[[i]] %>%
    full_join(y, by = c("geo_value", "time_value")) %>%
    mutate(geo_num = as.integer(as.factor(geo_value))) %>%
    group_by(geo_value) %>%
    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) +
             delta * (geo_num - 1)) %>%
    ungroup() %>%
    ggplot(aes(x = time_value, y = value, group = geo_value)) +
    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
    geom_line(alpha = 0.5, color = colors[i], size = line_width) +
    theme_void() +
    annotate("text", x = xp, y = yp, label = names[i],
             color = colors[i], angle = 90, size = label_size) +
    theme(panel.background = element_rect(fill = "black"))
}

gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
ggsave(file = "cover-graphic.pdf", gg, height = 22.5, width = 21, units = "cm")
```

### 2021-11-04 perspective/paper.tex
Order citations by appearance; use numbered style
Affiliations and reordering
Add Astley paragraph

\documentclass{article}
\input{ryantibs}

\title{Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year}
\author{Roni Rosenfeld \and Ryan J. Tibshirani}

@@

\item \emph{Authors' note to the Editor: this is to be filled out later.  The
    Astley et al. paper is still undergoing a significant revision and its main
    messages may change as a result.}
\end{enumerate}

@@

\newpage
\bibliographystyle{plainnat}
\bibliography{../common/covidcast}

>>>>

\documentclass{article}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}     % use 8-bit T1 fonts
\usepackage{booktabs}        % professional-quality tables
\usepackage{nicefrac}          % compact symbols for 1/2, etc.
\usepackage{microtype}       % microtypography
\usepackage{newtxtext}       % times font

\usepackage[numbers]{natbib}
\usepackage{amssymb,amsmath,amsthm,bbm}

\usepackage[margin=1in]{geometry}
\usepackage{verbatim,float,url,dsfont}
\usepackage{graphicx,subfigure,psfrag}
\usepackage{algorithm,algorithmic}
\usepackage{mathtools,enumitem}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}

\title{Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year}
\author{Roni Rosenfeld$^a$ \and Ryan J. Tibshirani$^{a,b}$}
\date{$^a$Machine Learning Department, Carnegie Mellon University \\ 
$^b$Department of Statistics \& Data Science, Carnegie Mellon University}

@@

\item \citet{Astley:2021} focuses on the international version of CTIS, which is
  an (ongoing) online survey operated by the University of Maryland, again in
  partnership with Facebook.  This international survey covers over 100
  countries, and is run in coordination with the U.S. one, so that they bear a
  similar structure and undergo similar updates; the full data set of
  individual, anonymized survey responses is again available to individual
  researchers under a data use agreement.  The paper presents analyses that
  reflect some basic and important characteristics of the international survey,
  reflecting its value abroad, where public health reporting efforts may be more
  limited than those in the U.S.
\end{enumerate}

@@

\newpage
\bibliographystyle{unsrt}
\bibliography{../common/covidcast}
